---
title: "Understanding Patterns of Road Safety"
author: "Yicong Wu, Yujun Wang, Sibo Wang"

output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: cerulean
    highlight: tango
---

### Preamble: Loading packages and data

```{r}
library(ggplot2)
library(ISLR)
library(MASS)
library(klaR)
library(knitr)
library(glmnet)
library(plyr)
library(gam)
library(cluster) 
library(fpc)
library(ggmap)
library(randomForest)
library(pROC)
library(rattle)
library(rpart.plot)
library(nnet)
library(leaps)
library(boot)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

options(scipen = 4)
```

For this problem we'll be working with two years of bikeshare data from the Capital Bikeshare system in Washington DC.  The dataset contains daily bikeshare counts, along with daily measurements on environmental and seasonal information that may affect the bikesharing.  

### Data pre-processing 

Let's start by loading the data.

```{r}
AccidentData <- read.csv("http://www.andrew.cmu.edu/user/achoulde/95791/projects/Project%20D/DfTRoadSafety_Accidents_2012.csv", header = TRUE)
```

Now I am going to split these data 


```{r}
set.seed(531)
days=rep(0,length(AccidentData$Time))

AccidentData$Day_of_Week[which(AccidentData$Day_of_Week==7)]=1
days[which(AccidentData$Day_of_Week==1)] = "Weekend"
days[which(AccidentData$Day_of_Week==7)] = "Weekend"
AccidentData$Day_of_Week[which(AccidentData$Day_of_Week!= 1)] = 0
days[which(AccidentData$Day_of_Week!= 1)] = "Workday"

timePeriod=rep(0,length(AccidentData$Time))
time=rep(0,length(AccidentData$Time))
timePeriod[which(as.numeric(AccidentData$Time)< 421-60)]="Night"      #1: 00:00-06:59   
time[which(as.numeric(AccidentData$Time)< 421-60)]=1
AccidentData$Time[which(as.numeric(AccidentData$Time)< 421-60)]="23:59"


timePeriod[which(as.numeric(AccidentData$Time)< 542+60)]="Morning Rush Hours"      #1: 07:00-09:00
time[which(as.numeric(AccidentData$Time)< 542+60)] = 2
AccidentData$Time[which(as.numeric(AccidentData$Time)< 542+60)]="23:59"

timePeriod[which(as.numeric(AccidentData$Time)< 1021-60)]="Day Time"      #1: 09:01-16:59
time[which(as.numeric(AccidentData$Time)< 1021-60)] = 3
AccidentData$Time[which(as.numeric(AccidentData$Time)< 1021-60)]="23:59"

timePeriod[which(as.numeric(AccidentData$Time)< 1142+60)]="Night Rush Hours"      #1: 17:00-19:00
time[which(as.numeric(AccidentData$Time)< 1142+60)] = 4
AccidentData$Time[which(as.numeric(AccidentData$Time)< 1141+60)]="23:59"

timePeriod[which(timePeriod == 0)]= "Night"
AccidentData$time[which(as.numeric(AccidentData$time)==0)] = 1





AS<- rep(0,length(AccidentData$Accident_Severity))
AS[which(AccidentData$Accident_Severity==1)] = "Fatal"
AS[which(AccidentData$Accident_Severity==2)] = "Serious"
AS[which(AccidentData$Accident_Severity==3)] = "Slight"


Roadtype=rep(0,length(AccidentData$Road_Type))

Roadtype[which(AccidentData$Road_Type==1)]="1"
Roadtype[which(AccidentData$Road_Type==2)]="2"
Roadtype[which(AccidentData$Road_Type==3)]="3"
Roadtype[which(AccidentData$Road_Type==4)]="4"
Roadtype[which(AccidentData$Road_Type==5)]="5"
Roadtype[which(AccidentData$Road_Type==6)]="6"
Roadtype[which(AccidentData$Road_Type==7)]="7"
Roadtype[which(AccidentData$Road_Type==8)]="8"
Roadtype[which(AccidentData$Road_Type==9)]="9"

Pedestrian.humancontrol=rep(0,length(AccidentData$Pedestrian_Crossing.Human_Control))

Pedestrian.humancontrol[which(AccidentData$Pedestrian_Crossing.Human_Control==0)]="0"
Pedestrian.humancontrol[which(AccidentData$Pedestrian_Crossing.Human_Control==1)]="1"
Pedestrian.humancontrol[which(AccidentData$Pedestrian_Crossing.Human_Control==2)]="2"

Lightcondition=rep(0,length(AccidentData$Light_Conditions))

Lightcondition[which(AccidentData$Light_Conditions==1)]="1"
Lightcondition[which(AccidentData$Light_Conditions==2)]="2"
Lightcondition[which(AccidentData$Light_Conditions==3)]="3"
Lightcondition[which(AccidentData$Light_Conditions==4)]="4"
Lightcondition[which(AccidentData$Light_Conditions==5)]="5"
Lightcondition[which(AccidentData$Light_Conditions==6)]="6"
Lightcondition[which(AccidentData$Light_Conditions==7)]="7"

Weathercondition=rep(0,length(AccidentData$Weather_Conditions))

Weathercondition[which(AccidentData$Weather_Conditions==1)]="1"
Weathercondition[which(AccidentData$Weather_Conditions==2)]="2"
Weathercondition[which(AccidentData$Weather_Conditions==3)]="3"
Weathercondition[which(AccidentData$Weather_Conditions==4)]="4"
Weathercondition[which(AccidentData$Weather_Conditions==5)]="5"
Weathercondition[which(AccidentData$Weather_Conditions==6)]="6"
Weathercondition[which(AccidentData$Weather_Conditions==7)]="7"
Weathercondition[which(AccidentData$Weather_Conditions==8)]="8"
Weathercondition[which(AccidentData$Weather_Conditions==9)]="9"

Roadsurfacecondition=rep(0,length(AccidentData$Road_Surface_Conditions))

Roadsurfacecondition[which(AccidentData$Road_Surface_Conditions==-1)]="-1"
Roadsurfacecondition[which(AccidentData$Road_Surface_Conditions==1)]="1"
Roadsurfacecondition[which(AccidentData$Road_Surface_Conditions==2)]="2"
Roadsurfacecondition[which(AccidentData$Road_Surface_Conditions==3)]="3"
Roadsurfacecondition[which(AccidentData$Road_Surface_Conditions==4)]="4"
Roadsurfacecondition[which(AccidentData$Road_Surface_Conditions==5)]="5"


UrbanorRural=rep(0,length(AccidentData$Urban_or_Rural_Area))

UrbanorRural[which(AccidentData$Urban_or_Rural_Area==1)]="1"
UrbanorRural[which(AccidentData$Urban_or_Rural_Area==2)]="2"

Speedlimit=rep(0,length(AccidentData$Speed_limit))

Speedlimit[which(AccidentData$Speed_limit==10)]="10"
Speedlimit[which(AccidentData$Speed_limit==20)]="20"
Speedlimit[which(AccidentData$Speed_limit==30)]="30"
Speedlimit[which(AccidentData$Speed_limit==40)]="40"
Speedlimit[which(AccidentData$Speed_limit==50)]="50"
Speedlimit[which(AccidentData$Speed_limit==60)]="60"
Speedlimit[which(AccidentData$Speed_limit==70)]="70"

AS2<- rep(0,length(AccidentData$Accident_Severity))
AS2[which(AccidentData$Accident_Severity==1)] = "Fatal"
AS2[which(AccidentData$Accident_Severity==2)] = "Fatal"
AS2[which(AccidentData$Accident_Severity==3)] = "Slight"


AccidentData <- data.frame(AccidentData,time,timePeriod,days,AS,Roadtype,Pedestrian.humancontrol,Lightcondition,Weathercondition,Roadsurfacecondition,UrbanorRural,Speedlimit,AS2)








# Randomly select 20% of the data to be held out for model validation
AccidentData.1.test.idx <- sample(which(AccidentData$Accident_Severity == 1),length(which(AccidentData$Accident_Severity == 1))*0.2,replace =TRUE)
AccidentData.2.test.idx <- sample(which(AccidentData$Accident_Severity == 2),length(which(AccidentData$Accident_Severity == 2))*0.2,replace =TRUE)
AccidentData.3.test.idx <- sample(which(AccidentData$Accident_Severity == 3),length(which(AccidentData$Accident_Severity == 3))*0.2,replace =TRUE)
AccidentData.test.ind <- c(AccidentData.1.test.idx,AccidentData.2.test.idx,AccidentData.3.test.idx)
AccidentData.train.ind <- setdiff(1:nrow(AccidentData), AccidentData.test.ind)

# Just pull the covariates available to marketers (cols 1:8) and the outcome (col 17)
AccidentData.train <- AccidentData[AccidentData.train.ind, c(11,33,17,18,23,25,26,27,30,7,34,35,36)]
AccidentData.test <- AccidentData[AccidentData.test.ind, c(11,33,17,18,23,25,26,27,30,7,34,35,36)]
AccidentData.location <- AccidentData[,c(4,5,7,13)]

AccidentData.train.nb <- AccidentData[AccidentData.train.ind, c(34,35,37,38,39,40,41,42,43,44)]
AccidentData.test.nb <- AccidentData[AccidentData.test.ind, c(34,35,37,38,39,40,41,42,43,44)]
```


```{r, cache = TRUE}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
     usr <- par("usr"); on.exit(par(usr))
     par(usr = c(0, 1, 0, 1))
     r <- abs(cor(x, y))
     txt <- format(c(r, 0.123456789), digits = digits)[1]
     txt <- paste0(prefix, txt)
     if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
     text(0.5, 0.5, txt, cex = pmax(1, cex.cor * r))
}
#pairs(~.,data=AccidentData.train.nb,lower.panel = panel.cor)
```
##Fit these data with multinomal regression.  And use Forward stepwise selection to select the best predictors.
Try to fit the data with multinomal regression.
```{r,cache =TRUE}

mu.fit <- multinom(Accident_Severity ~ .-timePeriod-days-AS, data = AccidentData.train)
#names(coef(fit.km))[which(coef(summary(fit))[,4]<=0.05)]
Accident.subset <- regsubsets(Accident_Severity ~ .-timePeriod-days-AS,
               data = AccidentData.train,
               nbest = 1,    # 1 best model for each number of predictors
               nvmax = NULL,    # NULL for no limit on number of variables
               method = "forward", really.big = TRUE)
print(names(coef(Accident.subset,id=5)))


```

Construct a plot showing R-squared on the y-axis and model size on the x-axis.
```{r}
qplot(x=1:9,y=summary(Accident.subset)$rsq,xlab="Model Size",ylab="R-squared",geom=c("line", "point"))
qplot(x=1:9,y=summary(Accident.subset)$cp,xlab="Model Size",ylab="AIC",geom=c("line", "point"))
qplot(x=1:9,y=summary(Accident.subset)$bic,xlab="Model Size",ylab="BIC",geom=c("line", "point"))
```
We can see from the plot, the R-squared are pretty small, So the prediction is pretty unreliable.


##Clustring the Longitude and Latitude

Run Clustering method on Longitude and Latitude

```{r,cache = TRUE}
#plot the longitude and latitude in 1-D
#plot(AccidentData.location$Longitude)






mapgilbert <- get_map(location = c(lon = mean(AccidentData.location$Longitude), lat = mean(AccidentData.location$Latitude)), zoom = 5, scale = 2)

ggmap(mapgilbert) + geom_point(data = AccidentData.location, aes(x = Longitude, y = Latitude, fill = AccidentData.location$Accident_Severity, alpha = 0.2), size = 2, shape = 21) +guides(fill=FALSE, alpha=FALSE, size=FALSE)
```




```{r}
#rescaling and plot
rateLetter <- c()
lon <- c()
lat <- c()
rate <-c()

priorRate <- length(which(AccidentData.location$Accident_Severity != 3))/length(AccidentData.location$Accident_Severity)
for( i in unique(AccidentData.location$Local_Authority_.District.)){
  
  r = length(which(AccidentData.location[which(AccidentData.location$Local_Authority_.District. == i),]$Accident_Severity != 3))/length(AccidentData.location[which(AccidentData.location$Local_Authority_.District. == i),]$Accident_Severity)
  rate <- c(rate,r)
  if(r>priorRate){
    rateLetter <- c(rateLetter,"High Serious Rate")
  }
  else{
    rateLetter <- c(rateLetter,"Low Serious Rate")
  }
  
  lon <- c(lon,mean(AccidentData.location[which(AccidentData.location$Local_Authority_.District. == i),]$Longitude))
  lat <- c(lat,mean(AccidentData.location[which(AccidentData.location$Local_Authority_.District. == i),]$Latitude))
}
geomInfo <- data.frame(rate,lon,lat,rateLetter)

ggmap(mapgilbert) + geom_point(data = geomInfo, aes(x = lon, y = lat, fill = rateLetter, alpha = 0.2), size = 2, shape = 21)

geomInfo.km <- data.frame(rate,lon,lat)
geomInfo.km.fit <- kmeans(geomInfo.km, 10)
geomInfo.km.fit.center <- data.frame(geomInfo.km.fit$center)
geomInfo.km.fit.center[,"rate"][which(geomInfo.km.fit.center[,"rate"]>priorRate)] <- "High Serious Rate"
geomInfo.km.fit.center[,"rate"][which(geomInfo.km.fit.center[,"rate"]<priorRate)] <- "Low Serious Rate"

ggmap(mapgilbert) + geom_point(data = geomInfo.km.fit.center, aes(x = lon, y = lat, fill = rate), size = 2, shape = 21)
```



```{r,cache = TRUE}
Accident.rf = randomForest(Accident_Severity~.-timePeriod-days-AS,data=AccidentData.train,mtry=4,importance=TRUE)
print (Accident.rf)
```

```{r}
importance(Accident.rf)
varImpPlot(Accident.rf)

```

```{r}
pred=predict(Accident.rf,AccidentData.test)
class = rep(0,length(pred))
class[which(pred >= 2.5)] = 3
class[which(pred>=1.5 && pred < 2.5)]= 2
class[which(pred<1.5)]=1
class
modelroc <- roc(AccidentData.test$Accident_Severity ,class)
plot(modelroc)

```




```{r,cache = TRUE}

#Accident.tr <- rpart(Accident_Severity ~.-timePeriod-days-AS, data = AccidentData.train, 
#                        control = rpart.control(minsplit=100, cp=0.002))
#plotcp(Accident.tr)
#Accident.tr$cptable
#Accident.pruned=prune(Accident.tr, cp = Accident.tr$cptable[2,1])
#plot(Accident.pruned)
#text(Accident.pruned)
```


These data fitting method shows poor prediction.



##Conditional probability.
Here we want to calculte some conditional probability. 

Here is some result using Naive Bayes. 
```{r}
# Naive Bayes 

accident.nb <- NaiveBayes(AccidentData.train.nb$AS2~.,data=AccidentData.train.nb,usekernel = TRUE)

nb.perdict=predict(accident.nb,newdata=AccidentData.test.nb)
# Confusion matrix

confusion.nb <- table(nb.perdict$class, AccidentData.test.nb$AS2)
confusion.nb

# Misclassification rate:
1 - sum(diag(confusion.nb)) / sum(confusion.nb)
```

We want to see the Likelyhood of Serious Accident happen given each condition. So we can tell in which condition the accident will be more serious. 

Here we conbined the Accident Severity 1 and 2 as serious, and leave 3 as slight. And in each plot, the horizontal line is the Prior of the serious accident happen. 
```{r,cache=TRUE}

PFatal <- length(which(AccidentData.train.nb$AS2=="Fatal"))/nrow(AccidentData.train.nb)
PSlight <- length(which(AccidentData.train.nb$AS2=="Slight"))/nrow(AccidentData.train.nb)

predit <- colnames(AccidentData.train.nb)
posteroriResult <- list()
for( i in predit[1:length(predit)-1]){
  resultif <- c()
  resultis <- c()
  CP <- eval(parse(text = paste("accident.nb$tables$",i)))
  
  type <- colnames(CP)
  for(j in type){
    Pij <-length(which(eval(parse(text = paste("AccidentData.train.nb$",i,sep="")))==j))/length(eval(parse(text = paste("AccidentData.train.nb$",i,sep=""))))
    
    
    resultif <- c(resultif,CP["Fatal",j]*PFatal/Pij)
    
    resultis <- c(resultis,CP["Slight",j]*PSlight/Pij)
  }
  
  assign( i,data.frame("Fatal" = resultif,"Slight"= resultis,row.names = type))
  #posteroriResult[length(posteroriResult)+1] <- eval(parse(text = i))
  
  posteroriResult[[paste(i)]]<- eval(parse(text = i))
  draw <- qplot( x=type,y=resultif,xlab = paste(i),ylab = paste("P(Fatal|",i,")",sep=""))+ geom_hline(yintercept=PFatal)

  print(draw)

}
print(posteroriResult)


```